---
title: "Social Web Analytics (Project)"
author: "Aadil Joshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
```

# 2.1 Data Collection: Anatomy of a Blockbuster

Data source: The Movie Database (TMDb) REST API (v3). Outputs: `movies.csv` (500 rows) and `cast.csv` (~5,000 rows).

First run only: `install.packages(c("httr", "jsonlite", "dplyr", "purrr", "readr"))`

```{r config}
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(readr)

API_KEY  <- "TMDB_API_KEY"
BASE_URL <- "https://api.themoviedb.org/3"


`%||%` <- function(a, b) if (is.null(a)) b else a

set.seed(3020)          # reproducible sampling
N_MOVIES  <- 500
N_HORROR  <- 60         # target (>= 50 required)
N_COMEDY  <- 60         # target (>= 50 required)
HORROR_ID <- 27
COMEDY_ID <- 35
```

```{r helper}
# One API call with retries
tmdb_get <- function(path, query = list()) {
  query$api_key <- API_KEY
  res <- RETRY("GET", paste0(BASE_URL, path), query = query,
               times = 5, pause_base = 1, pause_cap = 10, quiet = TRUE)
  stop_for_status(res)
  fromJSON(content(res, as = "text", encoding = "UTF-8"), simplifyVector = FALSE)
}

# Quick key check: fails immediately if the key is wrong
invisible(tmdb_get("/configuration"))
message("API key works.")
```

## Candidate pool via /discover/movie

Queried year by year (2015-2024) so we are not capped by TMDb's 500-page limit and every year is represented.

```{r pool}
discover_year <- function(year) {
  q <- list(
    language = "en-US",
    with_original_language = "en",
    `primary_release_date.gte` = paste0(year, "-01-01"),
    `primary_release_date.lte` = paste0(year, "-12-31"),
    `vote_count.gte` = 100,
    `with_runtime.gte` = 60,      
    include_adult = "false",
    include_video = "false",
    sort_by = "popularity.desc",
    page = 1
  )
  first <- tmdb_get("/discover/movie", q)
  pages <- min(first$total_pages, 500)
  message(year, ": ", first$total_results, " films, ", pages, " pages")

  map_dfr(seq_len(pages), function(p) {
    q$page <- p
    out <- if (p == 1) first else tmdb_get("/discover/movie", q)
    map_dfr(out$results, function(m) {
      tibble(
        id = m$id,
        title = m$title %||% NA_character_,
        overview = m$overview %||% NA_character_,
        release_date = m$release_date %||% NA_character_,
        genre_ids = paste(unlist(m$genre_ids), collapse = "|"),
        popularity = m$popularity %||% NA_real_,
        vote_average = m$vote_average %||% NA_real_,
        vote_count = m$vote_count %||% NA_integer_,
        original_language = m$original_language %||% NA_character_
      )
    })
  })
}

pool <- map_dfr(2015:2024, discover_year) %>%
  distinct(id, .keep_all = TRUE) %>%
  filter(!is.na(overview), trimws(overview) != "",
         vote_count >= 100, original_language == "en") %>%
  mutate(
    is_horror = grepl(paste0("(^|\\|)", HORROR_ID, "(\\||$)"), genre_ids),
    is_comedy = grepl(paste0("(^|\\|)", COMEDY_ID, "(\\||$)"), genre_ids),
    pop_tier  = ntile(popularity, 4)   # popularity quartiles
  )

message("Candidate pool: ", nrow(pool), " films (",
        sum(pool$is_horror), " horror, ", sum(pool$is_comedy), " comedy)")
stopifnot(sum(pool$is_horror) >= N_HORROR, sum(pool$is_comedy) >= N_COMEDY)
```

## Stratified sample across popularity levels

```{r sample}
take_strat <- function(df, n) {
  per <- ceiling(n / 4)
  df %>%
    group_by(pop_tier) %>%
    slice_sample(n = per) %>%
    ungroup() %>%
    slice_sample(n = min(n, nrow(.)))
}

horror <- take_strat(filter(pool, is_horror), N_HORROR)
comedy <- take_strat(filter(pool, is_comedy, !id %in% horror$id), N_COMEDY)
chosen <- bind_rows(horror, comedy)

rest_n <- N_MOVIES - nrow(chosen)
rest   <- take_strat(filter(pool, !id %in% chosen$id), rest_n)
movies <- bind_rows(chosen, rest) %>% distinct(id, .keep_all = TRUE)

# top up if rounding/duplicates left us short
while (nrow(movies) < N_MOVIES) {
  extra <- pool %>% filter(!id %in% movies$id) %>%
    slice_sample(n = N_MOVIES - nrow(movies))
  movies <- bind_rows(movies, extra)
}
movies <- movies %>% slice_head(n = N_MOVIES)

movies_out <- movies %>%
  select(id, title, overview, release_date, genre_ids, popularity,
         vote_average, vote_count, original_language)
```

## Top 10 credited actors per movie

```{r cast}
get_cast <- function(movie_id) {
  Sys.sleep(0.05)
  cr <- tryCatch(tmdb_get(paste0("/movie/", movie_id, "/credits")),
                 error = function(e) NULL)
  if (is.null(cr) || length(cr$cast) == 0) return(NULL)
  cast <- cr$cast
  ord  <- order(map_dbl(cast, ~ .x$order %||% Inf))
  cast <- cast[ord][seq_len(min(10, length(cast)))]
  tibble(
    movie_id   = movie_id,
    actor_id   = map_int(cast, ~ .x$id),
    actor_name = map_chr(cast, ~ .x$name)
  )
}

cast_out <- map_dfr(movies_out$id, get_cast)
```

## Integrity checks and save

```{r save}
stopifnot(
  nrow(movies_out) == N_MOVIES,
  !anyDuplicated(movies_out$id),
  all(trimws(movies_out$overview) != ""),
  all(cast_out$movie_id %in% movies_out$id)
)
message("Horror films: ", sum(grepl("(^|\\|)27(\\||$)", movies_out$genre_ids)))
message("Comedy films: ", sum(grepl("(^|\\|)35(\\||$)", movies_out$genre_ids)))
message("Cast rows: ", nrow(cast_out), " | movies with cast: ",
        n_distinct(cast_out$movie_id))
print(summary(movies_out[, c("popularity", "vote_average", "vote_count")]))

write_csv(movies_out, "movies.csv")
write_csv(cast_out,   "cast.csv")
message("Saved movies.csv and cast.csv")
```


# Read Files:
```{r}
movies <- read.csv("movies.csv", col_types = cols(genre_ids = col_character()))
```


# ---- 1. Define the two groups ------------------------------------------
```{r}
has_genre <- function(x, id) grepl(paste0("(^|\\|)", id, "(\\||$)"), x)
 
n_both <- sum(has_genre(movies$genre_ids, 27) & has_genre(movies$genre_ids, 35))
cat("Films tagged BOTH horror and comedy (excluded):", n_both, "\n")
 
dat <- movies %>%
  mutate(horror = has_genre(genre_ids, 27),
         comedy = has_genre(genre_ids, 35)) %>%
  filter(xor(horror, comedy)) %>%      # keep films in exactly one of the two groups
  mutate(genre = factor(if_else(horror, "Horror", "Comedy"),
                        levels = c("Horror", "Comedy")))
 
cat("\nGroup sizes:\n"); print(table(dat$genre))
if (any(table(dat$genre) < 50))
  warning("A group has fewer than 50 films - collect more horror/comedy films.")

```

